\textbf{Parallel dragonfly algorithm} \\[2ex] \large High Performance Computing for Data Science course 2024/2025

### 2025-02-11 text-paper/paper.tex
updated report

\documentclass[10pt]{article}
\usepackage{multicol}
\usepackage{authblk}
\usepackage{mathptmx}
\usepackage{graphicx}
\setlength{\columnsep}{5mm}
\usepackage{blindtext}
\usepackage{geometry}
\usepackage{url}

@@

\subsubsection*{How the algorithm works}
The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$; then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program. This procedure is simulated through the observation of repetitive patterns observed in the behaviour of dragonflies, which are:

@@

>>>>

\documentclass[10pt]{article}
\usepackage{multicol}
\usepackage{authblk}
\usepackage{mathptmx}
\usepackage{graphicx}
\setlength{\columnsep}{5mm}
\usepackage{blindtext}
\usepackage{geometry}
\usepackage{url}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{float}
\usepackage{longtable}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\small,
    columns=flexible,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=2pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

@@

\subsection*{How the algorithm works}
The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$; then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program. This procedure is simulated through the observation of repetitive patterns observed in the behaviour of dragonflies, which are:

@@

Now that we have seen the overall procedure it is possible to write a naive implementation of the algorithm in order to 
have a base for further developments. An example of the implementation can be observed in the following code snippet:

\begin{lstlisting}[language=C,caption={first implementation of the dragonfly algorithm}]
  void dragonfly_algorithm(Dragonfly *d,
                          float *average_speed, 
                          float *cumulated_pos, 
                          float * food, 
                          float * enemy, 
                          unsigned int N) {

    unsigned int dim = d->dim;
      // for each dragonfly
    for (unsigned int j = 0; j < d->N; j++) {
      float *cur_pos = d->positions + dim * j;
      float *cur_speed = d->speeds + dim * j;

      // compute separation: Si = -sumall(X-Xi)
      memcpy(d->S, cumulated_pos, sizeof(float) * dim);
      scalar_prod_assign(d->S, 1.0/(float)N, dim);
      sub_assign(d->S, cur_pos, dim);
      scalar_prod_assign(d->S, d->w.s, dim);

      // compute alignament: Ai = avarage(Vi)
      memcpy(d->A, average_speed, sizeof(float) * dim);
      scalar_prod_assign(d->A, d->w.a, dim);

      // compute cohesion: Ci = avarage(Xi)-X
      memcpy(d->C, cumulated_pos, sizeof(float) * dim);
      scalar_prod_assign(d->C, 1.0 / (float)N, dim);
      sub_assign(d->C, cur_pos, dim);
      scalar_prod_assign(d->C, d->w.c, dim);

      // food attraction: Fi=X_food - X
      memcpy(d->F, food, sizeof(float) * dim);
      sub_assign(d->F, cur_pos, dim);
      scalar_prod_assign(d->F, d->w.f, dim);

      // enemy repulsion: E=X_enemy+X
      memcpy(d->E, enemy, sizeof(float) * dim);
      sum_assign(d->E, cur_pos, dim);
      scalar_prod_assign(d->E, d->w.e, dim);

      brownian_motion(d->levy, dim, &d->seed);

      // compute speed = sSi + aAi + cCi + fFi + eEi + w
      scalar_prod_assign(cur_speed, d->w.w, dim);
      sum_assign(cur_speed, d->E, dim);
      sum_assign(cur_speed, d->F, dim);
      sum_assign(cur_speed, d->C, dim);
      sum_assign(cur_speed, d->A, dim);
      sum_assign(cur_speed, d->S, dim);
      sum_assign(cur_speed, d->levy, dim);

      // update current pos
      sum_assign(cur_pos, cur_speed, dim);
    }

    // update weights
    weights_step(&d->w);
  }
\end{lstlisting}

\noindent After the initialization of velocities and positions of agents,
which will be represented by two separated matricies with the same dimension equal to the number of members of our
population times the dimension of our problem, the algorithm starts to execute a for loop with a fixed amount of iteration stored in the variable $d\rightarrow N$, 
and will repeatetly compute the following steps:
\begin{itemize}
  \item retrieve current speed and position of the j-th dragonfly
  \item compute separation, alignment, cohesion, food attraction and enemy distraction vectors,
  stored into arrays placed in the struct Dragonfly
  \item calculate product between weigths and arrays
  \item update speed and position the dragonfly considering also levy flight
\end{itemize}
\subsection*{chunks}
We can improve this approach introducing groupments of dragonflies into chunks, 
allowing us to calculate local minima of different areas and compare them in order
to get as close as possible to the global minima. which will be considered
autonomous agents that will calculate local minima of our fitness function, updating then the 
global status of by sending messages to all other dragonfly chunks.

\begin{figure}[H]
  \includegraphics[scale=0.3]{chunks.png}
  \caption{illustration of communication between chunks}
\end{figure}

\section{Performace tests}

The serial implementation seen above can be easily tried out using some test function 
with known global minima in order to check the overall performances. We need to 
report time execution for each test function, using the input dimension as independent variable;
since however we have multiple input parameters to compare, we also have to decide which one will 
be considered during the performance evaluation.

\begin{table}[H]
\begin{tabular}{| c | c | c | c |}
  \hline
  \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
  \hline
  10 \\
  \hline

  100 \\
  \hline

  1000 \\
  \hline

  10000 \\ 
  \hline

  100000 \\
  \hline
\end{tabular}

\end{table}

\begin{table}[H]
  \begin{tabular}{| c | c | c | c |}
    \hline
    \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
    \hline
    10 \\
    \hline

    100 \\
    \hline

    1000 \\
    \hline

    10000 \\ 
    \hline

    100000 \\
    \hline
  \end{tabular}
\end{table}

\noindent As we can see the execution time increases exponentially with the consequentially increase of
the input.

\section{Parallel implementation}

\section{Parameters Retrievement}

\section{Speedup and final conclusions}
